---
title: "Test Linear SEM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TestLinearSEM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(wrapr)
#library(MASS)
library(Rcpp)
source("../R/CRT_statistic.R")
sourceCpp("../src/CRT_statistic.cpp")
source("../R/constraints_to_test.R")
source("../R/generate_data.R")
source("../R/Incomplete_U_statistic.R")
source("../R/dCovICA.R")
sourceCpp("../src/Incomplete_U_statistic.cpp")

# library(TestLinearSEM)
```

The package `TestLinearSEM` provides all the statistical tests that are used in the paper to test the null hypothesis if a distribution follows a linear structural equation model. For the details of the tests, see the paper. Here, you can find examples for using this package to apply the tests.

## Two nodes

1.  Generate data\
    First, we simulate data fulfilling the null hypothesis with number of nodes `p=2` and sample size `n=1000`.

```{r}
p <- 2 
l <- 0
n <- 100
data <- generate_data(n, p, l, "gamma", "H_0")
```

2.  Define constraints to test\
    Next, we define the matrix $M^{(3)}$ and the polynomial $\text{Str}$. Furthermore, we need all size $p+1$ minors of $M^{(3)}$.

```{r}
M <- construct_M(p, 2:3)
all_minors_M <-  calculate_minors(M, p+1) 
```

```{r}
strassen_ineq
```

3.  Statistical tests\
    CRT statistic:

```{r}
CRT(data, M, p)$PVAL
```

Incomplete $U$-statistic:

```{r}
incomplete_U_stat(data, equality_constraints = all_minors_M, ineq_constraints = list(strassen_ineq))$PVAL
```

Combination of incomplete $U$-statistic and CRT statistic:

```{r}
combine_pvalues = function(pval1, pval2) min(2*pval1, 2*pval2)

combine_pvalues(
      incomplete_U_stat(data, ineq_constraints = list(strassen_ineq))$PVAL,
      CRT(data, M, p)$PVAL)
```

Matteson-Tsay test:

```{r}
dCovICA(data)$PVAL
```

4.  Alternatives\
    The package also provides functionality to simulate data according to the alternatives presented in the Section 5 of the paper, namely the non-linear alternative and the latent confounding scenario:

```{r}
data <- generate_data(n, p, l, "gamma", "alt_cos", delta = 0.1)
data <- generate_data(n, p, l, "gamma", "alt_conf", delta = 0.1)
```

## Three nodes

```{r}
p <- 3
n <- 100
M <- construct_M(p, 2:3)

data <- generate_data(n, p, l, "gamma", "H_0")

CRT(data, M, p)$PVAL
```

```{r}
combine_pvalues(
  incomplete_U_stat(data, equality_constraints = list(Aronhold_invariant))$PVAL,
  CRT(data, M, p)$PVAL)
```

```{r}
dCovICA(data)$PVAL
```

## More nodes

```{r}
p <- 4
n <- 100
M <- construct_M(p, 2:3)
Y_T <- construct_Y_T(p, 3)
a <- floor((p-1)/2)
rank_Y_T <- p * choose(p-1, a)

data <- generate_data(n, p, l, "gamma", "H_0")

combine_pvalues(CRT(data, M, p)$PVAL, CRT(data, Y_T, rank_Y_T)$PVAL)
```
